A DPG method for compressible flow problems

### 2012-07-24 docs/Jesse/PECOStalk/proposalPres.tex
adding pecos talk

\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{color}
\usepackage{undertilde}
\usepackage{pdflscape}
\usepackage{pifont}

@@

\section{Proposed Work}
\frame{
\frametitle{DPG for nonlinear problems}
\textcolor{red}{REWORK. HOW TO DO THIS SECTION?}
Begin with the nonlinear variational formulation
$b(u,v) = l(v)$, linear in $v$, but not in $u$.  An appropriate linearization gives
\[
b_u(\Delta u,v) = l(v) - b(u,v),
\]
Let $B(u)$ and $B_u\Delta u$ be the variational operators associated with $b(u,v)$ and $b_u(\Delta u,v)$. 
\begin{align*}
\|\Delta u\|_E &\coloneqq \nor{B_u\Delta u}_{V'} = \nor{R_V^{-1} B_u\Delta u}_V \\
\nor{R(u)}_E &\coloneqq \nor{B(u)-l}_E = \nor{B(u) - l}_{V'}  = \nor{R_V^{-1} B(u)-l}_V 
\end{align*}
}
%\frame{
%\frametitle{Burgers' equation as an example}
%
%}
\frame{
\frametitle{Burgers equation}
\begin{columns}
\begin{column}{.48\textwidth}
\vspace{-.5cm}
\[
\pd{\left(u^2/2\right)}{x} + \pd{u}{y} + \epsilon \Delta u = f
\]
Burgers equation can be written with $\beta(u) = \LRp{u/2,1}$
\begin{align*}
\div\left(\beta(u)u-\sigma\right) &=f \\
\frac{1}{\epsilon}\sigma - \grad u &=0.
\end{align*}
i.e.\ nonlinear convection-diffusion.
\end{column}
\begin{column}{.48\textwidth}
\begin{figure}[!h]
\centering
\includegraphics[scale = .35]{figs/burgers1e4.png}
\caption{Shock solution for Burgers' equation, $\epsilon = 1e-4$.}
\end{figure}
\end{column}
\end{columns}
}

@@

\begin{figure}[!h]
\centering
\includegraphics[scale=.35]{figs/flat_plate_BCs.pdf}
\caption{Carter flat plate problem.}
\end{figure}
\begin{itemize}
\item{} Inflow and stress boundary conditions, momentum flux boundary conditions. 
\item{} No outflow condition set.
\end{itemize}
}
\frame{
\frametitle{Refinement level 0}
\begin{figure}
\subfigure[$u_1$]{\includegraphics[scale = .23]{figs/Re1000p2/u10.png}}
\subfigure[$T$]{\includegraphics[scale = .23]{figs/Re1000p2/T0.png}}
\subfigure{\includegraphics[scale = .35]{figs/Re1000p2/mesh0.png}}
\end{figure}
}
\frame{
\frametitle{Refinement level 1}
\begin{figure}
\subfigure[$u_1$]{\includegraphics[scale = .23]{figs/Re1000p2/u11.png}}
\subfigure[$T$]{\includegraphics[scale = .23]{figs/Re1000p2/T1.png}}
\subfigure{\includegraphics[scale = .35]{figs/Re1000p2/mesh1.png}}
\end{figure}
}
\frame{
\frametitle{Refinement level 2}
\begin{figure}
\subfigure[$u_1$]{\includegraphics[scale = .23]{figs/Re1000p2/u12.png}}
\subfigure[$T$]{\includegraphics[scale = .23]{figs/Re1000p2/T2.png}}
\subfigure{\includegraphics[scale = .35]{figs/Re1000p2/mesh2.png}}
\end{figure}
}
\frame{
\frametitle{Refinement level 3}
\begin{figure}
\subfigure[$u_1$]{\includegraphics[scale = .23]{figs/Re1000p2/u13.png}}
\subfigure[$T$]{\includegraphics[scale = .23]{figs/Re1000p2/T3.png}}
\subfigure{\includegraphics[scale = .35]{figs/Re1000p2/mesh3.png}}
\end{figure}
}
\frame{
\frametitle{Refinement level 4}
\begin{figure}
\subfigure[$u_1$]{\includegraphics[scale = .23]{figs/Re1000p2/u14.png}}
\subfigure[$T$]{\includegraphics[scale = .23]{figs/Re1000p2/T4.png}}
\subfigure{\includegraphics[scale = .35]{figs/Re1000p2/mesh4.png}}
\end{figure}
}
\frame{
\frametitle{Refinement level 5}
\begin{figure}
\subfigure[$u_1$]{\includegraphics[scale = .23]{figs/Re1000p2/u15.png}}
\subfigure[$T$]{\includegraphics[scale = .23]{figs/Re1000p2/T5.png}}
\subfigure{\includegraphics[scale = .35]{figs/Re1000p2/mesh5.png}}
\end{figure}
}
\frame{
\frametitle{Refinement level 6}
\begin{figure}
\subfigure[$u_1$]{\includegraphics[scale = .23]{figs/Re1000p2/u16.png}}
\subfigure[$T$]{\includegraphics[scale = .23]{figs/Re1000p2/T6.png}}
\subfigure{\includegraphics[scale = .35]{figs/Re1000p2/mesh6.png}}
\end{figure}
}

\section{Area A}
\frame{
\frametitle{Proposed work: Area A}
\begin{itemize}
\item{\textbf{Completed: Prove robustness of DPG method for the scalar convection-diffusion problem.}}

@@

\frame{  
\textbf{Proposed: a Nonlinear Hessian-based DPG method.}
Given a nonlinear variational problem $b(u,v) = l(v)$, linear in $v$ but not in $u$, beginning with the \textit{nonlinear} dual residual 
\[
J(u_h) = \frac{1}{2}\|B(u_h)-l\|_{V'}^2 \coloneqq\frac{1}{2} \sup_{v\in V\setminus\{0\}} \frac{| b(u_h,v)-l(v)|^2}{\nor{v}_V^2}.
\]
produces a Hessian-based DPG method 
\[
b_u(\Delta u,v) + b''(\Delta u, \delta u, v_{R(u)}) = r(u,v) = l(v) - b(u,v),
\]
which aims to minimize the nonlinear dual residual instead of the linearized problem residual. 
}

\section{Area C}
\frame{
\frametitle{Proposed work: Area C}\
textcolor{red}{GET GRAPHICS FOR EACH}
\begin{itemize}
\item{\textbf{Completed: convection-dominated diffusion, Burgers, and a model problem for Navier-Stokes.}} 
\item{\textbf{Proposed: Range of Reynolds/Mach numbers, ramp problem, Gaussian bump.}}
\item{\textbf{Proposed: regularized Euler.}}
\end{itemize}
}
\end{document}

>>>>

\usepackage{graphicx}
\usepackage{subfigure}
%\usepackage{caption}
%\usepackage{subcaption}

\usepackage{color}
\usepackage{undertilde}
\usepackage{pdflscape}
\usepackage{pifont}

@@

\section{Proposed Work}
\frame{
\frametitle{DPG for nonlinear problems}

Given some linearization technique, we can measure 
\begin{itemize}
\item the nonlinear residual 
\[
\nor{R(u)}_E \coloneqq \nor{B(u)-l}_E = \nor{B(u) - l}_{V'}  = \nor{R_V^{-1} B(u)-l}_V 
\]
\item size of the linearized update $\Delta u$
\[
\|\Delta u\|_E \coloneqq \nor{B_u\Delta u}_{V'} = \nor{R_V^{-1} B_u\Delta u}_V 
\]
\end{itemize}

%\textcolor{red}{REWORK. HOW TO DO THIS SECTION?}
%Begin with the nonlinear variational formulation
%$b(u,v) = l(v)$, linear in $v$, but not in $u$.  An appropriate linearization gives
%\[
%b_u(\Delta u,v) = l(v) - b(u,v),
%\]
%Let $B(u)$ and $B_u\Delta u$ be the variational operators associated with $b(u,v)$ and $b_u(\Delta u,v)$. 
%\begin{align*}
%\|\Delta u\|_E &\coloneqq \nor{B_u\Delta u}_{V'} = \nor{R_V^{-1} B_u\Delta u}_V \\
%\nor{R(u)}_E &\coloneqq \nor{B(u)-l}_E = \nor{B(u) - l}_{V'}  = \nor{R_V^{-1} B(u)-l}_V 
%\end{align*}
}
%\frame{
%\frametitle{Burgers' equation as an example}
%
%}
\frame{
\frametitle{Test case: Burgers equation}
\begin{columns}
\begin{column}{.48\textwidth}
\vspace{-.5cm}
\[
\pd{\left(u^2/2\right)}{x} + \pd{u}{y} + \epsilon \Delta u = f
\]
Burgers equation can be written with $\beta(u) = \LRp{u/2,1}$
\begin{align*}
\div\left(\beta(u)u-\sigma\right) &=f \\
\frac{1}{\epsilon}\sigma - \grad u &=0.
\end{align*}
i.e.\ nonlinear convection-diffusion.
\end{column}
\begin{column}{.48\textwidth}
\begin{figure}[!h]
\centering
\includegraphics[scale = .35]{figs/burgers1e4.png}
\caption{Shock solution for Burgers' equation, $\epsilon = 1e-4$.}
\end{figure}
\end{column}
\end{columns}
}

@@

\begin{figure}[!h]
\centering
\includegraphics[scale=.35]{figs/flat_plate_BCs.pdf}
\caption{Carter flat plate problem.}
\end{figure}
\begin{itemize}
\item{} Inflow and stress boundary conditions, momentum flux boundary conditions. 
\item{} No outflow condition set.
\end{itemize}
}
\frame{
\frametitle{Refinement level 0}
\begin{figure}
\subfigure[$u_1$]{\includegraphics[scale = .23]{figs/Re1000p2/u10.png}}
\subfigure[$T$]{\includegraphics[scale = .23]{figs/Re1000p2/T0.png}}
\subfigure{\includegraphics[scale = .35]{figs/Re1000p2/mesh0.png}}
\end{figure}
}
\setcounter{subfigure}{0}

\frame{
\frametitle{Refinement level 1}
\begin{figure}
\subfigure[$u_1$]{\includegraphics[scale = .23]{figs/Re1000p2/u11.png}}
\subfigure[$T$]{\includegraphics[scale = .23]{figs/Re1000p2/T1.png}}
\subfigure{\includegraphics[scale = .35]{figs/Re1000p2/mesh1.png}}
\end{figure}
}
\setcounter{subfigure}{0}

\frame{
\frametitle{Refinement level 2}
\begin{figure}
\subfigure[$u_1$]{\includegraphics[scale = .23]{figs/Re1000p2/u12.png}}
\subfigure[$T$]{\includegraphics[scale = .23]{figs/Re1000p2/T2.png}}
\subfigure{\includegraphics[scale = .35]{figs/Re1000p2/mesh2.png}}
\end{figure}
}
\setcounter{subfigure}{0}

\frame{
\frametitle{Refinement level 3}
\begin{figure}
\subfigure[$u_1$]{\includegraphics[scale = .23]{figs/Re1000p2/u13.png}}
\subfigure[$T$]{\includegraphics[scale = .23]{figs/Re1000p2/T3.png}}
\subfigure{\includegraphics[scale = .35]{figs/Re1000p2/mesh3.png}}
\end{figure}
}
\setcounter{subfigure}{0}

\frame{
\frametitle{Refinement level 4}
\begin{figure}
\subfigure[$u_1$]{\includegraphics[scale = .23]{figs/Re1000p2/u14.png}}
\subfigure[$T$]{\includegraphics[scale = .23]{figs/Re1000p2/T4.png}}
\subfigure{\includegraphics[scale = .35]{figs/Re1000p2/mesh4.png}}
\end{figure}
}
\setcounter{subfigure}{0}

\frame{
\frametitle{Refinement level 5}
\begin{figure}
\subfigure[$u_1$]{\includegraphics[scale = .23]{figs/Re1000p2/u15.png}}
\subfigure[$T$]{\includegraphics[scale = .23]{figs/Re1000p2/T5.png}}
\subfigure{\includegraphics[scale = .35]{figs/Re1000p2/mesh5.png}}
\end{figure}
}
\setcounter{subfigure}{0}

\frame{
\frametitle{Refinement level 6}
\begin{figure}
\subfigure[$u_1$]{\includegraphics[scale = .23]{figs/Re1000p2/u16.png}}
\subfigure[$T$]{\includegraphics[scale = .23]{figs/Re1000p2/T6.png}}
\subfigure{\includegraphics[scale = .35]{figs/Re1000p2/mesh6.png}}
\end{figure}
}
\setcounter{subfigure}{0}

\frame{
\frametitle{Zoomed solutions at the plate edge}
\begin{figure}
\subfigure[$\rho$]{\includegraphics[scale = .12]{figs/Re1000p2/rhoUnscaledzoom.png}}
\subfigure[$u_1$]{\includegraphics[scale = .12]{figs/Re1000p2/u1zoom.png}}
\subfigure[$u_2$]{\includegraphics[scale = .12]{figs/Re1000p2/u2zoom.png}}
\subfigure[$T$]{\includegraphics[scale = .12]{figs/Re1000p2/Tzoom.png}}
\subfigure[$q_n$]{\includegraphics[scale = .23]{figs/Re1000p2/heatflux.png}}
\end{figure}
}

\section{Proposed work}
\subsection{Area A}
\frame{
\frametitle{Proposed work: Area A}
\begin{itemize}
\item{\textbf{Completed: Prove robustness of DPG method for the scalar convection-diffusion problem.}}

@@

\frame{  
\textbf{Proposed: a Nonlinear Hessian-based DPG method.}
Given a nonlinear variational problem $b(u,v) = l(v)$, linear in $v$ but not in $u$, beginning with the \textit{nonlinear} dual residual 
\[
J(u_h) = \frac{1}{2}\|B(u_h)-l\|_{V'}^2 \coloneqq\frac{1}{2} \sup_{v\in V\setminus\{0\}} \frac{| b(u_h,v)-l(v)|^2}{\nor{v}_V^2}.
\]
produces a Hessian-based DPG method, which solves
\[
b_u(\Delta u,v) + b''(\Delta u, \delta u, v_{R(u)}) = l(v) - b(u,v) = r(u,v),
\]
and aims to minimize the nonlinear dual residual instead of the linearized problem residual. 
}

\section{Area C}
\frame{
\frametitle{Proposed work: Area C}
\begin{itemize}
\item{\textbf{Completed: convection-dominated diffusion, Burgers, and a model problem for Navier-Stokes.}} 
\item{\textbf{Proposed: Range of Reynolds/Mach numbers, ramp problem, Gaussian bump.}}
\item{\textbf{Proposed: regularized Euler.}}
\textcolor{red}{Graphics?}
\end{itemize}
}
\end{document}
